
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm, bm}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{physics}
\usepackage{bbm}

\title{Reality as a Self-Correcting Quantum Web:\\
A ToE Substrate via Tensor Networks, Quantum Error Correction, and QCA}
\author{}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We formalize a discrete quantum substrate in which reality is a network of finite-dimensional Hilbert spaces linked by entanglement. Local dynamics are unitary and strictly causal (QCA). Bulk-to-boundary holography arises as an isometric quantum error-correcting (QEC) map. Distances and geometry emerge from entanglement structure (minimal cuts); symmetries of the substrate induce gauge constraints in the bulk. We give axioms, operational definitions, and toy validations.
\end{abstract}

\section{Axioms (Formal)}
\begin{enumerate}
  \item \textbf{Discrete Substrate.} Let $G=(V,E)$ be a finite graph. Each $i\in V$ carries a Hilbert space $\mathcal{H}_i \cong \mathbb{C}^{d_i}$; the global space is $\mathcal{H} = \bigotimes_{i\in V}\mathcal{H}_i$.
  \item \textbf{Locality \& Causality.} Time evolution $U$ is a finite-depth circuit of local unitaries supported on neighborhoods in $G$. Thus there exists a finite Lieb--Robinson velocity $v_{LR}$.
  \item \textbf{Holographic Encoding.} There is an isometry $V:\mathcal{H}_{\mathrm{bulk}}\to \mathcal{H}_{\mathrm{bdy}}$ with QEC properties (Def.~\ref{def:KL}).
  \item \textbf{Error Robustness.} For an admissible family of errors $\{E_a\}$, Knill--Laflamme holds: $P E_a^\dagger E_b P = c_{ab} P$, $P=VV^\dagger$.
  \item \textbf{Emergent Geometry.} Entanglement induces a length functional; for tensor networks with bond dimension $\chi$, $S(A)\approx |\gamma_A| \log \chi$.
  \item \textbf{QCA Dynamics.} Discrete time: $\ket{\psi_{t+1}} = U \ket{\psi_t}$ with $U$ a QCA on $G$ (e.g., split-step).
  \item \textbf{Symmetry/Gauge Correspondence.} Global symmetries on the boundary act as local constraints (Gauss laws) in the bulk code subspace.
\end{enumerate}

\section{Operational Core}
\subsection{Substrate}
$G$ and local dimensions $\{d_i\}$ define the kinematics. Local neighborhoods $N_r(i)$ control admissible interactions.

\subsection{Quantum Error Correction}
\begin{defn}[Knill--Laflamme]\label{def:KL}
A subspace $\mathcal{C} = \mathrm{Im}(V)\subset \mathcal{H}$ corrects error set $\{E_a\}$ iff for all $a,b$,
$V^\dagger E_a^\dagger E_b V = c_{ab} \mathbbm{1}_{\mathrm{logical}}$.
\end{defn}

\subsection{Entanglement Geometry}
Define a cut functional $L(\gamma) = \sum_{e\in \gamma} \ell(e)$ with $\ell(e)=\log \chi_e$. For boundary region $A$, let $\gamma_A$ be a minimal cut homologous to $A$. Then
$S(A) \approx L(\gamma_A)$.

\subsection{QCA and Causal Cones}
Let $U=\prod_{\text{layers}} \prod_{(i,j)\in E_{\text{layer}}} U_{ij}$ be a checkerboard of local unitaries. Perturbations propagate within a light-cone bounded by depth $\times$ neighborhood radius.

\section{Toy Validations}
We include minimal checks: norm conservation under QCA, KL consistency on toy errors, and entanglement-cut proxies in 1D.

\bibliographystyle{plain}
\bibliography{refs}
\end{document}
